visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports

### 2024-07-29 manuscript.Rmd
updated manuscript

---
title             : "visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports"
shorttitle        : "VISUAL MGCFA"

author: 
  - name          : "Erin M. Buchanan"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "326 Market St., Harrisburg, PA, USA"
    email         : "ebuchanan@harrisburgu.edu"

affiliation:
  - id            : "1"
    institution   : "Harrisburg University of Science and Technology"

@@

abstract: |
  Latent variable modeling as a lens for psychometric theory is a popular tool for social scientists to examine measurement of constructs [@beaujean2014]. Journals such as *Assessment* regularly publish articles supporting measures of latent constructs wherein a measurement model is established. Confirmatory factor analysis can be used to investigate the replicability and generalizability of the measurement model in new samples, while multi-group confirmatory factor analysis is used to examine the measurement model across groups within samples [@brown2015]. With the rise of the replication crisis and "psychology's renaissance" [@nelson2018], interest in divergence in measurement has increased, often focused on small parameter differences within the latent model. This manuscript presents `visualizemi`, an *R* package that provides functionality to calculate multigroup models, partial invariance, visualizations for (non)-invariance, effect sizes for models and parameters, and potential replication rates compared to random models. Readers will learn how to interpret the impact and size of the proposed non-invariance in models with a focus on potential replication and how to plan for registered reports.

keywords          : "multigroup confirmatory factor analysis, measurement invariance, visualization, effect size"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
bibliography: includes/references.bib

@@

Measurement invariance has been explored and implemented for the last fifty years [@jöreskog1971; @sörbom1978] and implemented in the most popular structural equation modeling programs [@jöreskog2001; @rosseel2012; @boker2011]. @byrne1989 extended the ideas of multi-group testing by suggesting partial invariance [followed by @meredith1993]. Partial invariance occurs when non-invariance is found but can be attributed to only a few parameter estimate differences between groups (i.e., items 1 and 2 have different factor loadings but all others are the same). This testing provided an advantage to understand where the potential non-invariance may occur for further study and interpretation guidelines. As the field pushes back against favoring cutoff criteria and rules of thumb [@marsh2004; @putnick2016], an effect size measure for translating "how much" non-invariance was developed $d_{MACS}$ [@nye2011]. This effect size examines the differences in observed variables between the two groups for both the factor loading and the item intercept; thus, any differences in either or both will increase the effect size for non-invariance [@stark2006].

@@

Therefore, purpose of this manuscript is to describe an *R* package, `visualizemi`, that provides functionality to calculate multi-group confirmatory factor analysis, partial invariance tests, visualizations of the size of non-invariance, and potential effect sizes for overall models and individual parameters. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researcher's ability to judge the strength of the non-invariance differences and effect size for each item. The proposed effect sizes demonstrate the likelihood of replication with a similar sample as compared to a randomly assigned group model, thus, illustrating what type of measurement one might expect to find, and how different that is from random chance. Within this technique, the individual parameter effect sizes can be calculated: both the group differences within a model as compared to random and the likelihood of a parameter replication compared to random groups. Coupled with other indicators (i.e., fit indices differences, $d_{MACS}$), we can move toward a better understanding of how much measurement non-invariance is meaningful. This tutorial and package will help researchers plan future studies and aid in the ability to estimate a smallest effect of interest for measurement invariance studies, rather than relying on fit indices and rules of thumb alone.

@@

1.  Learn how to use *visualizemi* to analyze multi-group confirmatory factor analysis, examine partial for invariance, and create visualizations of parameters.
2.  Learn how to estimate the potential replication of multi-group models and their parameters using bootstrapping compared to a random group model.
3.  Be able to calculate and interpret effect sizes for model and parameter replication, as well as parameter group differences.
4.  Understand the impact of measurement variability on replication and generalizability.

The tutorial will start with simulated data based on known effect sizes using $d_{MACS}$ and demonstrating the package functions for 1) running the multigroup analysis, 2) running a partial invariance analysis, 3) plotting the partial invariance, 4) estimating replication and effect sizes at the model level, and 5) estimating replication and effect sizes at the parameter level. Last, data from @aiena2014 examining the measurement invariance of the RS-14 [@wagnild2009] will be used to demonstrate the application of the package on real data. The *visualizemi* package vignette includes an additional tutorial walk through.

# Method

## Design and Analysis

Data was simulated using the `simulateData` function in the *R* package *lavaan* [@rosseel2012] assuming multivariate normality using a $\mu$ of 0 and $\sigma$ of 1 for the data. This function allows you to write *lavaan* syntax for your model with estimated values to generate data for observed variables (see supplemental for examples). The data included two groups of individuals ("Group 1", "Group 2") for a multi-group confirmatory factor analysis ($n_{group}$ = 250, *N* = 500). The latent variables were assumed to be continuous normal (the package functions do not require this assumption). The model consisted of five observed items predicted by one latent variable (`lv =~ q1 + q2 + q3 + q4 + q5`); however, the demonstration in this manuscript extends to multiple latent variables and other combinations of observed variables. Each item was assumed to be related to the latent variable with loadings approximately equal to .40 to .80, except when cases of non-invariance on the loadings was simulated.

The @brown2015 steps of testing measurement invariance are demonstrated in this manuscript for illustration purposes, but in line with @stark2006 suggestions, the visualizations show the impact of loadings and intercepts together. A convenience function `mgcfa` is used for these steps or other measurement invariance test orders and combinations. Fit indices for the steps for multi-group models are presented in the appendix for comparison of cutoff rules of thumb [@cheung2002] to effect sizes and visualizations presented in this manuscript. Fit indices include Akaike Information Criterion [AIC, @akaike1998], Bayesian Information Criterion [BIC, @schwarz1978], Comparative Fit Index [CFI, @bentler1990], Tucker Lewis Index [TLI, @tucker1973], root mean squared error of approximation RMSEA [@steiger1990], and standardized root mean square residual [SRMR, @bentler1995].

The data was then simulated to represent invariance across all model steps, small, medium, and large invariance using $d_{MACS}$ estimated sizes from @nye2019. While $d_{MACS}$ is used primarily for an effect size of the (non)-invariance for intercepts and loadings together, a similar approach was taken for the estimation of small, medium, and large effects on the residuals. The effect size is presented for all models, calculated from the *dmacs* package [@dueber2023; @nye2011]. Only one item in each model was manipulated from the invariant model to create the non-invariant models. Given the data was simulated with a *z*-score scaling, the loading values were simulated at .30 points apart (given $d_{MACS}$ suggestions of .2, .4, .7), the intercepts at .25 points apart, and the residuals at .25 points apart. To plan a simulation for your own study, these values can be used to simulate small, medium, and large non-invariance effects by first converting data into *z*-score.

# Package Code Examples

The complete code for this manuscript can be found at <https://osf.io/wev5f/>. This tutorial was registered at <https://osf.io/vwf4d>, and the example provided at the end of the manuscript was added after that registration. The *R* package and replication/effect sizes was added after the original manuscript submission.

```{r create-models}
# overall model
model.overall <- "lv =~ q1 + q2 + q3 + q4 + q5
q1 ~ 0*1
lv ~ 1"

# build invariant model one factor
model.invariant.g1 <- "lv =~ .8*q1 + .4*q2 + .6*q3 + .3*q4 + .6*q5 
q4 ~~ 1*q4
q4 ~ 0*1
q1 ~ 0*1
lv ~ 1"
model.invariant.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ 1*q4
q4 ~ 0*1
q1 ~ 0*1
lv ~ 1"

# loadings .5, .8, 1.1
model.small.load.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .5*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"
model.med.load.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .8*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"
model.large.load.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + 1.1*q4 + .61*q5
q1 ~ 0*1
lv ~ 1"

# build invariance on intercepts .25, .50, and .75
model.small.int.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~ .25*1
q1 ~ 0*1
lv ~ 1"
model.med.int.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~ .5*1
q1 ~ 0*1
lv ~ 1"
model.large.int.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~ .75*1
q1 ~ 0*1
lv ~ 1"

# build invariance on residuals .25, .50, and .75
model.small.res.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ .75*q4
q1 ~ 0*1
lv ~ 1"
model.med.res.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ .5*q4
q1 ~ 0*1
lv ~ 1"
model.large.res.g2 <- "lv =~ .81*q1 + .41*q2 + .61*q3 + .31*q4 + .61*q5
q4 ~~ .25*q4
q1 ~ 0*1
lv ~ 1"
```

```{r create-data}
# simulate data invariant
df.invariant <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.invariant.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data small load
df.small.load <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.small.load.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data med load
df.med.load <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.med.load.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data large load
df.large.load <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.large.load.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data small int
df.small.int <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.small.int.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data med int
df.med.int <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.med.int.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data large int
df.large.int <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.large.int.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data small res
df.small.res <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.small.res.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data med res
df.med.res <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.med.res.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)

# simulate data large res
df.large.res <- bind_rows(
  simulateData(model.invariant.g1, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 1"), 
  simulateData(model.large.res.g2, sample.nobs = 250, 
               meanstructure = T, model.type = "cfa",
               seed = 1234) %>% 
    mutate(group = "Group 2") 
)
```

```{r run-cfa}
# invariant CFA
results.invariant <- 
  mgcfa(model = model.overall, 
        data = df.invariant,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

d_invariant <- lavaan_dmacs(
  results.invariant$model_configural, 
  RefGroup = "Group 1")

# loadings
results.small.load <- 
  mgcfa(model = model.overall, 
        data = df.small.load,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

results.med.load <- 
  mgcfa(model = model.overall, 
        data = df.med.load,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)
results.large.load <- 
  mgcfa(model = model.overall, 
        data = df.large.load,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

d_small_load <- 
  lavaan_dmacs(results.small.load$model_configural, 
               RefGroup = "Group 1")
d_med_load <- 
  lavaan_dmacs(results.med.load$model_configural, 
               RefGroup = "Group 1")
d_large_load <- 
  lavaan_dmacs(results.large.load$model_configural, 
               RefGroup = "Group 1")

# intercepts
results.small.int <- 
  mgcfa(model = model.overall, 
        data = df.small.int,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)
results.med.int <- 
  mgcfa(model = model.overall, 
        data = df.med.int,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)
results.large.int <- 
  mgcfa(model = model.overall, 
        data = df.large.int,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

d_small_int <- 
  lavaan_dmacs(results.small.int$model_configural, 
               RefGroup = "Group 1")
d_med_int <- 
  lavaan_dmacs(results.med.int$model_configural, 
               RefGroup = "Group 1")
d_large_int <- 
  lavaan_dmacs(results.large.int$model_configural, 
               RefGroup = "Group 1")

# residuals
results.small.res <-
  mgcfa(model = model.overall, 
        data = df.small.res,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)
results.med.res <- 
  mgcfa(model = model.overall, 
        data = df.med.res,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)
results.large.res <- 
  mgcfa(model = model.overall, 
        data = df.large.res,
        group = "group",
        group.equal = c("loadings", "intercepts", "residuals"),
        meanstructure = T)

d_small_res <- 
  lavaan_dmacs(results.small.res$model_configural, 
               RefGroup = "Group 1")
d_med_res <- 
  lavaan_dmacs(results.med.res$model_configural,
               RefGroup = "Group 1")
d_large_res <- 
  lavaan_dmacs(results.large.res$model_configural, 
               RefGroup = "Group 1")
```

## Multi-group CFA Caculation

First, we would create our model code in *lavaan* syntax [@rosseel2012]. The `lv` latent variable predicts the five measured variables, which are present as columns in our `df.invariant` data set. The package does generally require raw data for bootstrapping purposes, and an example of how to simulate data from models and covariance/correlations tables sometimes provided in manuscripts (rather than the raw data) is provided in the supplemental documentation.

*lavaan* automatically sets the mean (i.e., the intercept) for latent variables to zero. If we wish to visualize the impact of the changes in parameter estimates across groups on the latent means, we need to allow the latent mean estimation with `lv ~ 1`. However, adding this estimation into our model will create a non-identified model. To solve this problem, you can set one of the intercepts of another variable to a value to scale the model. Here we will set the scale of the model by using `q1 ~ 0*1`, thus, scaling the expected means to zero. With simulation, this step is easy to know which variable to pick - we set the intercept on the variable we know did not show differences. In real data, you may wish to run the model steps *without* setting this option, examine the results of a configural or separate models, and then add the option for the values most similar. Additionally, you could complete partial invariance steps to determine which value appears most consistent to fix the estimate.

```{r model-example, eval = T, include = T, echo = T, size = "small"}
# create lavaan model
model.overall <- "
# overall one-factor model
lv =~ q1 + q2 + q3 + q4 + q5
# set the intercept (mean) of q1 to zero
q1 ~ 0*1
# allow the lv intercept to be freely estimated
lv ~ 1"
# look at the data
head(df.invariant)
```

The `mgcfa` function is designed to flexibly allow you to leverage `lavaan`'s package functions to calculate multiple measurement steps at once. You would include:

1)  the model syntax in the `model` argument
2)  the dataframe in the `data` argument of our function
3)  the name of the grouping variable in quotes for `group`
4)  and the equality constraints you would like to impose in order in `group.equal`
5)  `...` any other *lavaan* arguments you would like to use such as `meanstructure` or `estimator`.

Note: you can also use `sample.cov`, `sample.mean`, `sample.nobs` in this step for estimation of multi-group models, but simulated dataframes are needed for bootstrapping replication estimates.

```{r mgcfa-example, echo = T, size = "small"}
# run our mgcfa function to run all models
results.invariant <- 
  # name of the saved model syntax
  mgcfa(model = model.overall, 
        # name of the dataframe
        data = df.invariant,
        # name of the grouping variable
        group = "group",
        # equality constraints to impose in order
        group.equal = c("loadings", "intercepts", "residuals"),
        # other options to send to lavaan cfa function
        meanstructure = T)

# what is saved for you
names(results.invariant)
```

@@

1)  `model_coef`: The parameter estimates for each model with the model step included in a *model* column. This set of coefficients can be used for other functions. This dataframe is created with *broom*'s `tidy()` function if you wish to recreate this table without running the `mgcfa()` function [@robinson2023].

```{r mgcfa-output1, echo = T, include = T, size = "small"}
results.invariant$model_coef[1:10 , ]
```

2)  `model_fit`: The model fit indices from `fitmeasures()` to review for overall model fit and invariance judgments. The name of the model is included in a *model* column.

```{r mgcfa-output2, echo = T, include = T, size = "small"}
head(results.invariant$model_fit)
```

3)  `model_overall`: A saved *lavaan* fitted model of all groups together without any equality constraints or grouping variables. These objects can be used with any function that normally takes a saved model: `parameterEstimates()`, `modificationIndices()`, `semPlot::semPaths()`, and so on [@epskamp2022].
4)  `group_models`: A list of saved fitted models for each group separately.
5)  `model_configural`: A saved fitted model for the configural model that nests together each group into one model with no other constraints.
6)  `invariance_models`: A list of saved fitted models that consecutively adds `group.equal` constraints.

## Visualization of Invariance

### Package Function

The results from the `model_coef` table can then be used directly in `plot_mi()`. The plot outputs will be described below. First, here are the arguments for the function:

1)  `data_coef`: A tidy dataframe of the parameter estimates from the models. This function assumes you have used `broom::tidy()` on the saved model from *lavaan* and added a column called "model" with the name of the model step [@robinson2023]. This function will only run for models that have used the grouping function (i.e., configural, metric, scalar, and strict or other combinations/steps you wish to examine).
2)  `model_step`: Which model do you want to plot? You should match this name to the one you want to extract from your model column in the `data_coef`.
3)  `item_name`: Which observed variable from your model syntax do you want to plot? Please list this variable name exactly how it appears in the model.
4)  `x_limits`: What do you want the x-axis limits to be for your invariance plot? The default option is to assume the latent variable is standardized, and therefore, -1 to 1 is recommended. Use only two numbers, a lower and upper limit. This value also constrains the latent mean diagram to help zoom in on group differences because the scale of latent means is usually centered over zero. You can use this parameter to zoom out to a more traditional histogram using `c(-2, 2)`.
5)  `y_limits`: What do you want the y-axis limits to be for your invariance plot? Given that the latent variable is used to predict the observed values in the data, you could use the minimum and maximum values found in the data. If that range is large, consider reducing this value to be able to visualize the results (i.e., otherwise it may be too zoomed out to judge group differences). Use only two numbers, a lower and upper limit.
6)  `conf.level`: What confidence limit do you want to plot? Use 1 - $\alpha$.
7)  `model_results`: In this argument, include the saved *lavaan* output for the model listed in the `model_step` argument.
8)  `lv_name`: Include the name of the latent variable, exactly how it is listed in your *lavaan* syntax. You should plot the latent variable that the `item_name` is linked to. If you have items that load onto multiple latent variables, you will need to make multiple plots.
9)  `plot_groups`: If you include more than two groups in a multi-group model, the automatic assumption is that you want the first two groups for this visualization. If not, include the names of the groups here to plot.

```{r plot-example, eval = T, echo = T, include = T, size = "small"}
invariant.plot <- 
  plot_mi(
    # output from model_coef
    data_coef = results.invariant$model_coef, 
    # which model do you want to plot
    model_step = "Configural", 
    # name of observed item
    item_name = "q4", 
    # latent variable limits to graph
    x_limits = c(-1,1), 
    # Y min and max in data 
    y_limits = c(min(df.invariant$q4), max(df.invariant$q4)),
    # what ci do you want
    conf.level = .95, 
    # what model results do you want 
    model_results = results.invariant$model_configural,
    # which latent variable do you want 
    lv_name = "lv" 
)

names(invariant.plot)
```

@@

### Simulated Results

>>>>

---
title: "visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports"
shorttitle: "VISUAL MGCFA"

author: 
  - name: "Erin M. Buchanan"
    affiliation: "1"
    corresponding: yes    # Define only one corresponding author
    address: "326 Market St., Harrisburg, PA, USA"
    email: "ebuchanan@harrisburgu.edu"

affiliation:
  - id: "1"
    institution: "Harrisburg University of Science and Technology"

@@

abstract: |
  Latent variable modeling as a lens for psychometric theory is a popular tool for social scientists to examine measurement of constructs [@beaujean2014]. Journals such as *Assessment* regularly publish articles supporting measures of latent constructs wherein a measurement model is established. Confirmatory factor analysis can be used to investigate the replicability and generalizability of the measurement model in new samples, while multigroup confirmatory factor analysis is used to examine the measurement model across groups within samples [@brown2015]. With the rise of the replication crisis and "psychology's renaissance" [@nelson2018], interest in divergence in measurement has increased, often focused on small parameter differences within the latent model. This manuscript presents `visualizemi`, an *R* package that provides functionality to calculate multigroup models, partial invariance, visualizations for (non)-invariance, effect sizes for models and parameters, and potential replication rates compared to random models. Readers will learn how to interpret the impact and size of the proposed non-invariance in models with a focus on potential replication and how to plan for registered reports.

keywords: "multigroup confirmatory factor analysis, measurement invariance, visualization, effect size"

floatsintext: no
linenumbers: yes
draft: no
mask: no

figurelist: no
tablelist: no
footnotelist: no

classoption: "man"
bibliography: includes/references.bib

@@

Measurement invariance has been explored and implemented for the last fifty years [@jöreskog1971; @sörbom1978] and implemented in the most popular structural equation modeling programs [@jöreskog2001; @rosseel2012; @boker2011]. @byrne1989 extended the ideas of multigroup testing by suggesting partial invariance [followed by @meredith1993]. Partial invariance occurs when non-invariance is found but can be attributed to only a few parameter estimate differences between groups (i.e., items 1 and 2 have different factor loadings but all others are the same). This testing provided an advantage to understand where the potential non-invariance may occur for further study and interpretation guidelines. As the field pushes back against favoring cutoff criteria and rules of thumb [@marsh2004; @putnick2016], an effect size measure for translating "how much" non-invariance was developed $d_{MACS}$ [@nye2011]. This effect size examines the differences in observed variables between the two groups for both the factor loading and the item intercept; thus, any differences in either or both will increase the effect size for non-invariance [@stark2006].

@@

Therefore, purpose of this manuscript is to describe an *R* package, `visualizemi`, that provides functionality to calculate multigroup confirmatory factor analysis, partial invariance tests, visualizations of the size of non-invariance, and potential effect sizes for overall models and individual parameters. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researcher's ability to judge the strength of the non-invariance differences and effect size for each item. The proposed effect sizes demonstrate the likelihood of replication with a similar sample as compared to a randomly assigned group model, thus, illustrating what type of measurement one might expect to find, and how different that is from random chance. Within this technique, the individual parameter effect sizes can be calculated: both the group differences within a model as compared to random and the likelihood of a parameter replication compared to random groups. Coupled with other indicators (i.e., fit indices differences, $d_{MACS}$), we can move toward a better understanding of how much measurement non-invariance is meaningful. This tutorial and package will help researchers plan future studies and aid in the ability to estimate a smallest effect of interest for measurement invariance studies, rather than relying on fit indices and rules of thumb alone.

@@

1.  Learn how to use *visualizemi* to analyze multigroup confirmatory factor analysis, examine partial for invariance, and create visualizations of parameters.
2.  Learn how to estimate the potential replication of multigroup models and their parameters using bootstrapping compared to a random group model.
3.  Be able to calculate and interpret effect sizes for model and parameter replication, as well as parameter group differences.
4.  Understand the impact of measurement variability on replication and generalizability.

The tutorial will start with a general overview of relevant topics to orient readers to invariance testing and MGCFA effect sizes. Next, the reader will learn about the `visualizemi` package functions for 1) running the multigroup analysis, 2) running a partial invariance analysis, 3) plotting the partial invariance, 4) estimating replication and effect sizes at the model level, and 5) estimating replication and effect sizes at the parameter level. Last, data from @aiena2014 and @chen2020 examining the measurement invariance of the RS-14 [@wagnild2009] will be used to demonstrate the application of the package on real data. The *visualizemi* package vignette includes an additional tutorial walk through.

# Terminology

## MGCFA

Multigroup confirmatory factor analysis (MGCFA) was proposed as a method to examine differences in scale functioning across groups [@brown2015] using structural equation modeling and confirmatory factor analysis (CFA). The goal of MGCFA is to determine if groups are invariant or show the same response patterns on the scale. MGCFA is often performed in sequential steps to determine the location and impact of potential non-invariance. The most common procedure starts by examining overall scale structure for each group individually to show that the groups can be combined into one nested model (i.e., both models at least converge). The combined model, often described as the *configural* model, creates one CFA for both groups that allows each group's estimated parameters to vary. Equality constraints between group parameters are then added sequentially to the model [@brown2015]. If parameters are found to be equivalent between groups, these models are considered "invariant", suggesting that any subtle differences in the parameter estimate should not effect overall scoring and assessment.

The first equality constraint added is usually the item factor loadings (*metric* model) which requires items to be related to the latent variable at the same strength across groups. The intercepts are then constrained to be equal across groups (*scalar* model) to determine the invariance of item averages. The item variances (*strict* model) can then be examined to determine if the general variation in item answers are equal across groups. Other parameter equality constraints can be set at the latent variable level (latent means, variance, covariances), but the focus is generally on the observable variables and their parameter estimations. Non-invariance is determined by examining differences in fit indices between models with constraints versus without [@cheung2002].

## Partial Invariance

If a specific step within the MGCFA testing framework shows non-invariance, partial invariance is often used to investigate where and "how much" invariance occurs [@meredith1993; @byrne1989]. Similar to *post hoc* follow up tests in ANOVA, each equality constraint for that model is examined one at a time by allowing groups to vary. If the model fit improves when groups are allowed to have separate parameter estimations, the item parameter estimate is considered non-invariant. When only a few parameters are found that impact invariance, models are considered partially invariant, implying that most, but not all parameters are equal across groups. The researcher then interprets the impact of those items and parameters on overall scores and assessment usefulness. Partial invariance investigation is a useful tool for finding specific items that vary between groups, but does not fully explain the effect size of the difference between groups and the impact on the overall model.

## Effect Size: $d_{MACS}$

$d_{MACS}$ was developed from Differential Item Functioning (DIF) measurement in Item Response Theory [IRT, @stark2004] wherein the effect size (DIF) portrays the group differences that lead to overall item score differences. Item scores can be mathematically defined as:

$$
\hat{X}_{iR} = \tau_{iR} + \lambda_{iR}\xi
$$

*i* indicates an individual item, and *R* indicates the reference group [@nye2011]. Therefore, an individual score is predicted by the intercept of the item ($\tau$) added to the item loading ($\lambda$) times the latent variable score ($\xi$). @nye2011 demonstrated that the terminology from IRT and DIF can be used to create a measure of item functioning differences within CFA, $d_{MACS}$. $d_{MACS}$ is then the difference of reference group versus focal group (e.g., group 1 versus group 2) divided by a pooled standard deviation similar to formulas for *d* proposed by @cohen2013. Therefore, if $\tau$ or $\lambda$ are individually (or both) different across groups, it will impact all the predicted $X_{iR}$ scores, and thus, impact $d_{MACS}$ values. While $d_{MACS}$ represents a necessary step for development of effect sizes within CFA, it does not separate the differences in parameters between groups in a way that can be paired with traditional MGCFA testing and partial invariance. In addition to the missing effect sizes at model and parameter levels, no effect size to date gives the researcher a feeling for potential replication of the invariant or non-invariant items. Last, in line with general suggestions by @cumming2012 and @cumming2016, the visualization of effect sizes in MGCFA would be an added tool for researchers to gauge the size of group differences.

# Package Functions

The %>% code for this manuscript can be found at <https://osf.io/wev5f/>. This tutorial was registered at <https://osf.io/vwf4d>, and the example provided at the end of the manuscript was added after that registration. The *R* package and replication/effect sizes was added after the original manuscript submission. The simulation study used to design plotting functions and test effect sizes can be found in the supplemental materials, along with worked code examples.

## MGCFA: `mgcfa()` Function

First, we would create our model code in *lavaan* syntax [@rosseel2012]. The *visualizemi* package does generally require raw data for bootstrapping purposes, and an example of how to simulate data from models and covariance/correlations tables that sometimes are provided in manuscripts (rather than the raw data) is provided in the supplemental documentation. The `mgcfa()` function is designed to flexibly allow you to leverage *lavaan*'s package functions to calculate multiple measurement steps at once. You would include:

1)  the model syntax in the `model` argument.
2)  the dataframe in the `data` argument of our function
3)  the name of the grouping variable in quotes for `group`.
4)  and the equality constraints you would like to impose in order in `group.equal`.
5)  `...` any other *lavaan* arguments you would like to use such as `meanstructure` or `estimator`.

@@

1)  `model_coef`: The parameter estimates for each model with the model step included in a *model* column. This set of coefficients can be used for other functions. This dataframe is created with *broom*'s `tidy()` function if you wish to recreate this table without running the `mgcfa()` function [@robinson2023].
2)  `model_fit`: The model fit indices from `fitmeasures()` to review for overall model fit and invariance judgments. The name of the model is included in a *model* column.
3)  `model_overall`: A saved *lavaan* fitted model of all groups together without any equality constraints or grouping variables. These objects can be used with any function that normally takes a saved model: `parameterEstimates()`, `modificationIndices()`, `semPlot::semPaths()`, and so on [@epskamp2022].
4)  `group_models`: A list of saved fitted models for each group separately.
5)  `model_configural`: A saved fitted model for the configural model that nests together each group into one model with no other constraints.
6)  `invariance_models`: A list of saved fitted models that consecutively adds `group.equal` constraints.

## Partial Invariance: `partial_mi()` Function

The `partial_mi()` function aids in the calculation of partial invariance for a specific step of the MGCFA process. The function includes the following arguments:

1)  `saved_model`: The saved *lavaan* model with the equality constraints at the level of measurement invariance you would like to examine for partial invariance.
2)  `data`: The dataframe where the model was estimated.
3)  `model`: The model syntax for the overall model.
4)  `group`: The grouping variable column in the dataframe.
5)  `group.equal`: The equality constraints including in your original multigroup tests.
6)  `partial_step`: The level of partial invariance you wish to test.

In this function, each parameter with the appropriate *lavaan* syntax is relaxed individually (i.e., `~1` for intercepts, `~~` for residuals, etc.). The fitted models are saved in the `models` output, and the `fit_table` output includes all fit indices for each model to investigate potential areas of partial invariance based on the researcher's desired criterion.

## Visualization of Invariance: `plot_mi()` Function

Once we know which items are non-invariant, the `model_coef` output from the `mgcfa()` can be used directly in `plot_mi()`. The plot outputs will be described below. First, here are the arguments for the function:

1)  `data_coef`: A tidy dataframe of the parameter estimates from the models. This function assumes you have used `broom::tidy()` on the saved model from *lavaan* and added a column called "model" with the name of the model step [@robinson2023]. This function will only run for models that have used the grouping function (i.e., configural, metric, scalar, and strict or other combinations/steps you wish to examine).
2)  `model_step`: Which model do you want to plot? You should match this name to the one you want to extract from your model column in the `data_coef`.
3)  `item_name`: Which observed variable from your model syntax do you want to plot? Please list this variable name exactly how it appears in the model.
4)  `x_limits`: What do you want the x-axis limits to be for your invariance plot? The default option is to assume the latent variable is standardized, and therefore, -1 to 1 is recommended. Use only two numbers, a lower and upper limit. This value also constrains the latent mean diagram to help zoom in on group differences because the scale of latent means is usually centered over zero. You can use this parameter to zoom out to a more traditional histogram using `c(-2, 2)`.
5)  `y_limits`: What do you want the y-axis limits to be for your invariance plot? Given that the latent variable is used to predict the observed values in the data, you could use the minimum and maximum values found in the data. If that range is large, consider reducing this value to be able to visualize the results (i.e., otherwise it may be too zoomed out to judge group differences). Use only two numbers, a lower and upper limit.
6)  `conf.level`: What confidence limit do you want to plot? Use 1 - $\alpha$.
7)  `model_results`: In this argument, include the saved *lavaan* output for the model listed in the `model_step` argument.
8)  `lv_name`: Include the name of the latent variable, exactly how it is listed in your *lavaan* syntax. You should plot the latent variable that the `item_name` is linked to. If you have items that load onto multiple latent variables, you will need to make multiple plots.
9)  `plot_groups`: If you include more than two groups in a multigroup model, the automatic assumption is that you want the first two groups for this visualization. If not, include the names of the groups here to plot.

@@

## Parameter Replication and Effect Sizes: `bootstrap_partial()` Function
